\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}
\usepackage{pdfsync}
\usepackage{enumitem}

\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{xspace}

\usepackage{siunitx}

\usepackage{floatflt}

\usepackage{multirow}

\usepackage{booktabs}

%\renewcommand{\baselinestretch}{2.0}
%\usepackage{lineno}
%\renewcommand\linenumberfont{\normalfont\tiny}
%\linenumbers

\graphicspath{{svg/}}

\usepackage[normalem]{ulem}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{todonotes}

\pgfplotsset{compat=1.9}
\definecolor{gnuplot@lightblue}{RGB}{87,181,232}
\definecolor{gnuplot@green}{RGB}{0,158,115}
\definecolor{gnuplot@purple}{RGB}{148,0,212}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\snes}{{\specialword{SNES}}\xspace}
\newcommand{\ts}{{\specialword{TS}}\xspace}
\newcommand{\petscsf}{{\specialword{SF}}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\candi}{{\specialword{candi}}\xspace}
\newcommand{\MPI}{{\specialword{MPI}}\xspace}
\newcommand{\MPIx}{{\specialword{MPI+X}}\xspace}
\newcommand{\sundials}{{\specialword{SUNDIALS}}\xspace}
\newcommand{\kinsol}{{\specialword{KINSOL}}\xspace}
\newcommand{\ida}{{\specialword{IDA}}\xspace}
\newcommand{\arkode}{{\specialword{ARKODE}}\xspace}
\newcommand{\boost}{{\specialword{Boost}}\xspace}
\newcommand{\kokkos}{{\specialword{Kokkos}}\xspace}
\newcommand{\llvm}{{\specialword{LLVM}}\xspace}
\newcommand{\step}[1]{{\specialword{step-#1}}\xspace}

%Trilinos Packages
\newcommand{\epetra}{{\specialword{Epetra}}\xspace}
\newcommand{\petra}{{\specialword{Petra}}\xspace}
\newcommand{\teuchos}{{\specialword{Teuchos}}\xspace}
\newcommand{\tpetra}{{\specialword{Tpetra}}\xspace}


\usetikzlibrary{shapes.misc}
\tikzset{cross/.style={cross out, draw=black, minimum size=2*(#1-\pgflinewidth), inner sep=0pt, outer sep=0pt},
%default radius will be 1pt.
cross/.default={2pt}}

%
% Author list -- please add yourself in both places below (in
%                alphabetical order) if you think that your
%                contributions to the last release warrant this
%

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Bruno Blais,
    Marc Fehling,
    Rene Gassm\"{o}ller,
    Timo Heister,
    Luca Heltai,
    Vladimir Ivannikov,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.8},
}

\title{The \dealii Library, Version 9.8}

 \author[1*]{Daniel Arndt}
 \affil[1]{Computational Coupled Physics Group,
   Computational Sciences and Engineering Division,
   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
   TN 37831, USA.
   \texttt{arndtd/turcksinbr@ornl.gov}}

 \author[2,3]{Wolfgang~Bangerth}
 \affil[2]{Department of Mathematics, Colorado State University, Fort
   Collins, CO 80523, USA.
   \texttt{bangerth/sam.scheuerman@colostate.edu}}
 \affil[3]{Department of Geosciences, Colorado State University, Fort
   Collins, CO 80523, USA.}

\author[5]{Bruno Blais}
\affil[5]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
             Polytechnique Montréal,
             PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
             {\texttt{bruno.blais@polymtl.ca}}}

\author[6]{Marc~Fehling}
\affil[6]{Department of Mathematical Analysis,
    Faculty of Mathematics and Physics, Charles University,
    Sokolovsk{\'a} 49/83, 186\,75 Prague 8, Czech Republic.
    {\texttt{marc.fehling@matfyz.cuni.cz}}}


\author[7]{Rene~Gassm\"{o}ller}
\affil[7]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany.
   {\texttt{rgassmoeller@geomar.de}}}

\author[8]{Timo~Heister}
\affil[8]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA.
   {\texttt{heister@clemson.edu}}}

\author[9]{Luca~Heltai}
\affil[9]{Department of Mathematics, University of Pisa,
Via Buonarroti 1/c, 56127 Pisa, Italy. {\texttt{luca.heltai@unipi.it}}}

\author[10]{Vladimir~Ivannikov}
\affil[10]{Institute of Material Systems Modeling, Helmholtz-Zentrum Hereon,
Max-Planck-Stra{\ss}e 1, 21502 Geesthacht, Germany. {\texttt{vladimir.ivannikov@hereon.de}}}

\author[11]{Martin~Kronbichler}
\affil[11]{Faculty of Mathematics, Ruhr University Bochum,
   Universit\"atsstr.~150, 44780 Bochum, Germany.
 {\texttt{martin.kronbichler@rub.de}}}

\author[12]{Matthias~Maier}
\affil[12]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[13]{Peter Munch}
\affil[13]{Institute of Mathematics, Technical University of Berlin, Germany.
  {\texttt{muench@math.tu-berlin.de}}}

\author[1*]{Bruno~Turcksin}

\author[15]{David Wells}
\affil[15]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\renewcommand{\labelitemi}{--}


\begin{document}
\maketitle

\footnotetext{%
  $^\ast$ This manuscript has been authored by UT-Battelle, LLC under Contract No.
  DE-AC05-00OR22725 with the U.S. Department of Energy.
  % We need to submit the manuscript with the text below. If the editor
  % complains we can remove it.
  The United States
  Government retains and the publisher, by accepting the article for
  publication, acknowledges that the United States Government retains a
  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
  the published form of this manuscript, or allow others to do so, for United
  States Government purposes. The Department of Energy will provide public
  access to these results of federally sponsored research in accordance with the
  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).
}


\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.8.
\end{abstract}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\dealii version 9.8.0 was released XXX.
\todo{Complete}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii software library version 9.8. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
terms of the \emph{GNU Lesser General Public License} (LGPL). The \dealii
project is in the process of relicensing the library under the terms of
the \emph{Apache License 2.0 with LLVM Exception}.
\todo[inline]{Do we need to update the license blurb?}
Downloads are
available at \url{https://www.dealii.org/} and
\url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
\item ...
  \todo[inline]{Complete}
\end{itemize}
%



While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
\item ...
  \todo[inline]{Complete}
\end{itemize}
%
The
\href{https://dealii.org/current/doxygen/deal.II/changes_between_9_7_0_and_9_8_0.html}{changelog}
-- listing more than XXX
new features and bugfixes --
contains a complete record of all changes; see \cite{changes98}.
\todo[inline]{Complete and check link.}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Major changes to the library}
\label{sec:major}

This release of \dealii contains a number of large and significant changes,
which we will discuss in this section.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Support for meshes with non-hypercube cells}
\label{subsec:non-hypercube}

\todo[inline]{Write about all of the stuff that extends
  pyramid/wedges/simplex support}

\subsection{Performance Portable GPU support}
\label{subsec:gpu}

\todo[inline]{Write Timo/Ryan, reference step-99 (if merged) and step-104}

\begin{itemize}
 \item Portable::MatrixFree: support for system of equations
 \item improved interface (for each quad point, DeviceVector, DeviceBlockVector)
 \item multi-threaded setup()
 \item improved documentation
 \item peformance improvements (reduced number of fences etc)
 \item support for asserts, error checks
 \item tensor and symmetric tensor support
 \item require Kokkos 4.0 for device support

\end{itemize}

\subsection{Trilinos and Tpetra interfaces}
\label{subsec:tpetra}

\todo[inline]{Write}
\texttt{Trilinos} has been talking about their \texttt{Epetra}-based solver stack in favor of \texttt{Tpetra}-based solver stack for a long time.
Where support for different number types and \texttt{Kokkos} execution spaces has been the main driver. Nevertheless, the \texttt{Epetra}-based stack
has only been removed in the recent \texttt{Trilinos} 17 release. 
While \texttt{TpetraWrappers::Vector} has already been implemented in the 9.1.0 \dealii release, the development for a full \texttt{Tpetra} has been slow since.
A basic implementation for \texttt{TpetraWrappers::SparsityPattern} and \texttt{TpetraWrappers::SparseMatrix} was included in the 9.6.0 \dealii release.
The release of the \texttt{Trilinos} 17 helped with making significant progress within this release cycle.
With this release, \texttt{Trilinos} support in \dealii only requires the Epetra-based stack or the Tpetra-based stack to be configured which in particular allows
support for Trilinos 17. In case \texttt{Epetra} is available, the \texttt{TrilinosWrappers} namespace continues to use \texttt{Epetra} but will use \texttt{Tpetra}
if only \texttt{Tpetra} is available. The \texttt{TpetraWrappers} namespace while always use \texttt{Tpetra}. While we generally expect the \texttt{Tpetra}
stack to work, not all functionality from the \texttt{TrilinosWrappers} namespace is (yet) available in the \texttt{TpetraWrappers} namespace. 



\subsection{VTK interfaces}
\label{subsec:vtk}

\todo[inline]{Write}


\subsection{Interfaces to SUNDIALS' ARKode package}
\label{subsec:arkode}

\todo[inline]{Write}


\subsection{xxx}
\label{subsec:xxx}

\todo[inline]{Add more sections}

\subsection{Reducing dynamic memory allocations}
\label{subsec:memory}

\todo[inline]{describe @drwell's work towards reducing dynamic memory
  allocations; mention the inplace vector with a reference to the Section~\ref{subsec:cxx}.}


\subsection{Work towards more modern C++ standards}
\label{subsec:cxx}

As in most previous releases, we have spent substantial work on
evolving our code base to keep track with developments in C++ and the
standards that define C++.

Specifically, we have continued our work on building C++20-style
modules (see the discussion in the previous release paper
\cite{dealii97} and the preprint
\cite{bangerth2025experienceconvertinglargemathematical}). Recent
compiler versions have become stricter at enforcing requirements
spelled out in the standard for how the source files for modules have to be
structured, and the scripts we use to convert the \dealii{} source
files to a module structure have correspondingly evolved. We have also
substantially expanded the module wrappers for the external projects
we interface with. Note that \dealii{} continues to only require
C++17, but can build C++20-style modules if so instructed and if the
compiler supports it.

Over the past year, the standardization committee has also released
the C++26 specifications. Among its additions is a class
\texttt{std::inplace\_vector<T,N>} that has a size that can vary at
runtime, but can not exceed \texttt{N}, allowing for efficient
allocation of such objects on the stack. In contrast, the widely used
class \texttt{std::vector<T>} has to allocate memory on the heap
dynamically, whereas \texttt{std::array<T,N>} has a fixed size
\texttt{N}. The new class is useful in contexts where the actual size
of the object is not known at compile time, but where it is known that
the size can not exceed a certain number; an example is the number of
faces per cell that depends on what kind of cell we have (say, a
tetrahedron or a hexahedron) but that in no case can exceed six.

We have written a fallback version of this class that we can use even
if the compiler in use does not support C++26. We have started to use
it in many places where the maximal size of a array is known at
compile time, and this has dramatically reduced the number of dynamic
memory allocations in many cases -- see also the discussion in
Section~\ref{subsec:memory}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New and improved tutorials and code gallery programs}
\label{subsec:steps}

Many of the \dealii tutorial programs were revised in a variety of ways
as part of this release: Around 125 of the more than 1500 (non-merge)
commits that went into this release touched the tutorial.
% data generated using these commands:
% - tutorial: git log --since 2024/08/11 --until 2025/07/19 --no-merges examples | grep commit | wc -l
% - total:    git log --since 2024/08/11 --until 2025/07/19 --no-merges | grep commit | wc -l
\todo[inline]{Update numbers}
In addition, there are a number of new tutorial
programs:
\begin{itemize}
\item
  \step{98} shows how to solve two-dimensional magnetostatic curl-curl
  problems.
    \step{98} was written by ...
    \todo[inline]{Expand and Complete}
  \item \step{100} applies a discontinuous Petrov-Galerkin method to the time-harmonic Helmholtz equation.
    \step{100} was written by ....
    \todo[inline]{Expand and Complete}
  \item \step{104} solves the Stokes problem on GPUs
    \step{104} was written by Quang and Timo
    \todo[inline]{Expand and Complete}
\end{itemize}

There are also several new programs in the code gallery (a collection of
user-contributed programs that often solve more complicated problems
than tutorial programs, and that are intended as starting points for further
research rather than as teaching tools):
\begin{itemize}
\item ``\textit{Parallel implementation of heat equation}'' is a transient heat
  equation solver with adaptive mesh refinement and a parallel solver;
  this program was written by XXX.
\item
  ``\textit{L-BFGS phasefield solver}'' is a limited-memory BFGS
  monolithic solver for phase-field crack simulations; it was written
  by XXX.
\item
  ``\textit{Agglomeration-based solver for the Poisson problem}'' is a
  discontinuous Galerkin solver for the Poisson problem on general
  polytopal meshes generated through mesh agglomeration;
\item
  ``\textit{An ALE approach for large-deformation thermoplasticity}''
  is an implementation of a large-deformation arbitrary
  Lagrangian-Eulerian finite-strain thermoplasticity solver;
\item
  ``\textit{Parallel flow routing}'' is a parallel solver for
  determining how much water is available at every point of a
  landscape; the program was contributed by Wolfgang Bangerth.
\item
  ``\textit{Parallel Vibroacoustic Solver} is a parallel
  frequency-domain vibroacoustic solver calculating the sound
  transmission loss of a concrete wall;
\item
  ``\textit{Multilevel Monte Carlo for random Darcy flow}'' is an implementation of a multilevel Monte Carlo method for random Darcy flow.
  \todo[inline]{Cross-check the exact name of these programs. Also
    list their authors.}
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Incompatible changes}\label{subsec:deprecated}

The 9.8 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_7_0_and_9_8_0.html}
     {around 25 incompatible changes};
     \todo{Update this number}
see \cite{changes98}. Many of these
incompatibilities remove previously deprecated functions or classes,
or require now widely available but newer versions of external
dependencies than previous \dealii{} versions. Others
change internal
interfaces that are not usually used in external
applications. The following are worth mentioning since they
are more broadly visible:
\begin{itemize}
\item ...
  \todo[inline]{Complete and update numbers above. also check links}
\end{itemize}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii}
\label{sec:cite}

\todo[inline]{Double check this section}

In order to justify the work the developers of \dealii put into this
software, we ask that papers using the library reference one of the
\dealii papers. This helps us justify the effort we put into this library.

There are various ways to reference \dealii. To acknowledge the use of
the current version of the library, \textbf{please reference the present
  document}. For up-to-date information and a bibtex entry
see
\begin{center}
  \url{https://www.dealii.org/publications.html}
\end{center}

The original \dealii paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}, and a more recent
publication documenting \dealii's design decisions is available as \cite{dealII2020design}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{multicols}{2}
  \vspace*{-36pt}
  \begin{itemize}[leftmargin=4mm]
    \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019, munch2022gc};
    \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007,fehlingbangerth2023};
    \item For partition-of-unity (PUM) and finite element enrichment methods:
           \cite{Davydov2016};
    \item For matrix-free and fast assembly techniques:
          \cite{KronbichlerKormann2012,KronbichlerKormann2019};
    \item For computations on lower-dimensional manifolds:
          \cite{DeSimoneHeltaiManigrasso2009};
    \item For curved geometry representations and manifolds:
          \cite{HeltaiBangerthKronbichlerMola2019};
    \item For integration with CAD files and tools:
          \cite{HeltaiMola2015};
    \item For boundary element computations:
          \cite{GiulianiMolaHeltai-2018-a};
    \item For the \texttt{LinearOperator} and
      \texttt{Packaged\-Operation} facilities:
          \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
    \item For uses of the \texttt{WorkStream} interface:
          \cite{TKB16};
    \item For uses of the \texttt{ParameterAcceptor} concept, the
          \texttt{MeshWorker::ScratchData} base class, and the
          \texttt{ParsedConvergenceTable} class:
          \cite{SartoriGiulianiBardelloni-2018-a};
    \item For uses of the particle functionality in \dealii:
      \cite{GLHPB18}.
      \item For the design of the video lectures and how they can be
        used in teaching: \cite{Zarestky2022}.
          \vfill\null
  \end{itemize}
\end{multicols}

\dealii can interface with many other libraries:
\todo{Do we need to update/add anything here?}
\begin{multicols}{4}
  \begin{itemize}[leftmargin=4mm]
    \item ADOL-C \cite{griewank1996adolc}
    \item ArborX \cite{prokopenko2025}
    \item ARPACK \cite{lehoucq1998arpack}
    \item Assimp \cite{schulze2021assimp}
    \item BLAS, LAPACK \cite{anderson1999lapack}
    \item Boost \cite{boost-web-page}
    \item CGAL \cite{cgal:eb-24b}
    \item Gmsh \cite{geuzaine2009gmsh}
    \item GSL \cite{galassi2009gsl,gsl-web-page}
    \item Ginkgo \cite{anzt2020ginkgo,anzt2022ginkgo}
    \item HDF5 \cite{hdf5-web-page}
    \item Kokkos \cite{trott2022}
    \item Magic Enum \cite{magic-enum-web-page}
    \item METIS \cite{karypis1998metis}
    \item MUMPS \cite{amestoy2001mumps,amestoy2019mumps}
    \item muparser \cite{muparser-web-page}
    \item OpenCASCADE \cite{opencascade-web-page}
    \item p4est \cite{burstedde2011p4est}
    \item PETSc \cite{petsc-user-ref,petsc-web-page}
    \item PSBLAS \cite{Filippone2000psblas}
    \item ROL \cite{ROL2022ICCOPT}
    \item ScaLAPACK \cite{blackford1997scalapack}
    \item SLEPc \cite{roman2023improvements}
    \item SUNDIALS \cite{gardner2022enabling}
    \item SymEngine \cite{symengine-web-page}
    \item Taskflow \cite{huang2021taskflow}
    \item TBB \cite{reinders2007tbb}
    \item Trilinos \cite{mayr2025trilinos,trilinos-web-page}
    \item UMFPACK \cite{davis2004umfpack}
    \item VTK \cite{vtk-web-page}
    \item zlib \cite{zlib-web-page}
  \end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use
interfaces to these libraries.

The two previous releases of \dealii can be cited as
\cite{dealii96,dealii97}.


\section{Acknowledgments}

\dealii is a worldwide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\

\todo[inline]{Create list and update text below.}

\begin{quote}
%  authors removed:
%  Daniel       Arndt,
%  Wolfgang     Bangerth,
%  Maximilian   Bergbauer,
%  Bruno        Blais,
%  Marc         Fehling,
%  Rene         Gassmoeller,
%  Timo         Heister,
%  Luca         Heltai,
%  Martin       Kronbichler,
%  Matthias     Maier,
%  Peter        Munch,
%  Jean-Paul    Pelteret,
%  Sam          Scheuerman,
%  Bruno        Turcksin,
%  Siarhei      Uzunbajakau,
%  David        Wells,
%  Michał       Wichrowski,
Pasquale     Africa,
Henry        Arhin,
Alfredo      Buttari,
Bruna        Campos,
Nicholas     Cantrell,
Xiaoming     Cao,
Chayapol     Chaoveeraprasit,
Jerett       Cherry,
Dario        Coscia,
John         Coughlin,
Nikita       Daniliuk,
Crystal      Farris,
Marco        Feder,
Emmanuel     Ferdman,
Federico     Fernandez,
Olivier      Gaboriault,
Mohamad      Ghadban,
Mahdi        Gharehbaygloo,
Robin        Görmer,
Davit        Gyulamiryan,
Lóránt       Hadnagy,
Fernando     Herrera,
Robin        Hiniborch,
Quang        Hoang,
Jordan       Hoffart,
Sascha       Hofstetter,
Yimin        Jin,
Yann         Jobic,
Sean         Johnson,
Vaishnavi    Kale,
Sebastian    Kinnewig,
Andreas      Koch,
Jason        Landini,
Zhou         Lei,
Yingli       Li,
Nils         Margenberg,
Oreste       Marquis,
Ryan         Moulday,
Nils         Much,
Tileuzhan    Mukhamet,
Wasim~Niyaz  Munshi,
Natalia      Nebulishvili,
Luz          Paz,
David        Pecoraro,
Davide       Polverino,
Sanjeeb      Poudel,
Guilhem      Poy,
Laura        Prieto~Saavedra,
Sebastian    Proell,
Andreas      Ritthaler,
Mayank       Sabharwal,
Magdalena    Schreter,
Richard      Schussnig,
Kyle         Schwiebert,
Marc         Secanell,
Qingyuan     Shi,
Wyatt        Smith,
Simon        Sticko,
Dominik      Still,
Edward       Terrell,
Jan~Philipp  Thiele,
Xiaochuan    Tian,
Mikael       Vaillant,
Vinayak      Vijay,
Stephan      Voss,
Simon        Wiesheier,
Yi-Yung      Yang.
\end{quote}
Their contributions are much appreciated -- this project lives by its
community of users and contributors!

This release contains patches written in and contributed from
Antarctica. As a consequence, \dealii{} now consists of work written
on all seven continents.

\bigskip

\dealii and its developers are financially supported through a
variety of funding sources:

\todo[inline]{Update as appropriate}

D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
LLC, for the U. S. Department of Energy and supported by the U.S. Department of Energy,
Office of Science, Office of Advanced Scientific Computing Research,
Next-Generation Scientific Software Technologies program, under contract
number DE-AC05-00OR22725.

W.~Bangerth was partially supported by the National Science Foundation
under awards OAC-1835673, EAR-1925595, and OAC-2410847.

W.~Bangerth, T.~Heister, and R.~Gassm\"{o}ller were partially
supported by the Computational Infrastructure for Geodynamics initiative
(CIG), through the National Science Foundation (NSF) under Award
No.~EAR-2149126 via The University of California -- Davis.

B.~Blais was supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant and the MMIAOW Canada Research Level 2 in Computer-Assisted Design and Scale-up of Alternative Energy Vectors for Sustainable Chemical Processes.

M.~Fehling was partially supported by the ERC-CZ grant LL2105 CONTACT,
funded by the Czech Ministry of Education, Youth and Sports. He was also
partially supported by the Charles University Research Centre Program No.\@ UNCE/24/SCI/005.

R.~Gassm\"{o}ller was also partially supported by NSF Awards EAR-1925677
and EAR-2054605.

T.~Heister was also partially supported by NSF Awards OAC-2015848, EAR-1925575, and OAC-2410848.

L.~Heltai is a member of Gruppo Nazionale per il Calcolo Scientifico (GNCS) of
Istituto Nazionale di Alta Matematica (INdAM). LH was partially supported by
the Italian Ministry of University and Research (MUR), under the grant MUR PRIN
2022 No. 2022WKWZA8 ``Immersed methods for multiscale and multiphysics problems
(IMMEDIATE)'', and acknowledges the MIUR Excellence Department Project awarded
to the Department of Mathematics, University of Pisa, CUP I57G22000700001.

M.~Kronbichler was partially supported by the
German Federal Ministry of Research, Technology and Space, project
``PDExa: Optimized software methods for solving partial differential
equations on exascale supercomputers'', grant agreement No. 16ME0637K.

M.~Kronbichler and L.~Heltai were partially supported by the European Research
Council (ERC) under the European Union's Horizon 2020 research and innovation
programme (call HORIZON-EUROHPC-JU-2023-COE-03, grant agreement No. 101172493
``dealii-X: an Exascale Framework for Digital Twins of the Human Body'').

M.~Maier was partially supported by NSF Award DMS-2045636 and by the
Air Force Office of Scientific Research under grant/contract number
FA9550-23-1-0007.

D.~Wells was supported by NSF Award OAC-1931516.

Charles University is acknowledged for providing computing time on the Sn\v{e}hurka cluster.

This research used in part resources on the Palmetto Cluster at
Clemson University under National Science Foundation awards MRI
1228312, II NEW 1405767, MRI 1725573, and MRI 2018069. The views
expressed in this article do not necessarily represent the views of
NSF or the United States government.
\todo[inline]{Timo: update as necessary}

\bibliography{paper}{}
\bibliographystyle{abbrv}

\end{document}
